---
title: 'Week 11: Machine Learning'
author: "Yograj Karki"
date: "6/5/2021"
output: pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages and working directories
```{r message= F, warning= F}
#Packages
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse)
library(factoextra)
library(cluster)
library(caret)
library(class)

  setwd("~/MSDS/DSC520/dsc520")
```

## Loading the data
```{r}
binary_df <- read.csv("data/binary-classifier-data.csv")
trinary_df <- read.csv("data/trinary-classifier-data.csv")

binary_df$label <- as.factor(binary_df$label)
trinary_df$label <- as.factor(trinary_df$label)

head(binary_df)
head(trinary_df)
```


# Scatter plot binary classifier
```{r}

library(ggvis)
binary_df %>% ggvis(~x, ~y, fill = ~label) %>% layer_points()

```

# Scatter plot trinary classifier data
```{r}
trinary_df %>% ggvis(~x, ~y, fill = ~label) %>% layer_points()

```


# k nearest neighbors algorithm
## Data preparation
```{r}
# Generate a random number that is 80% of the total number of rows in data set.
set.seed(42)
# Random sampling
random_binary <- sample(1:nrow(binary_df), 0.8 * nrow(binary_df))
random_trinary <- sample(1:nrow(trinary_df), 0.8*nrow(trinary_df))
```


## Splitting training and testing data
```{r}
#Binary
train_binary <- binary_df[random_binary,]
test_binary <- binary_df[-random_binary,]
#Trinary
train_trinary <- trinary_df[random_trinary,] 
test_trinary <- trinary_df[-random_trinary,] 


# Creating separate dataframe for our target
train_binary_labels <- binary_df[random_binary,1]
test_binary_labels <-binary_df[-random_binary,1]
train_trinary_labels <- trinary_df[random_trinary,1]
test_trinary_labels <-trinary_df[-random_trinary,1]
```

## Fit a k nearest neighbors’ model for each dataset for k=3, k=5, k=10, k=15, k=20, and k=25. Compute the accuracy of the resulting models for each value of k. Plot the results in a graph where the x-axis is the different values of k and the y-axis is the accuracy of the model.

## Confusion matrix formula

```{r}
# Formula for Binary confusion matrix
my.statistics <- function(Actual,Predicted) {
  confusion.table <- table(Actual=Actual,Predicted=Predicted)
  output <- list(confusion.table=confusion.table)
  TN <- confusion.table[1]
  FN <- confusion.table[2]
  FP <- confusion.table[3]
  TP <- confusion.table[4]
  output$accuracy <- (TP+TN)/sum(confusion.table)
  output$precision <- (TP)/(TP+FP)
  output$sensitivity <- (TP)/(TP+FN)
  output$specificity <- (TN)/(TN+FP)
  output$FPR <- (FP)/(TN+FP)
  
  return(output)
}
```


```{r}
# k = 3 on Binary data
knn_pred <- knn(train_binary[,c("x","y")],
                test_binary[,c("x", "y")],
                cl = train_binary$label, 
                k = 15)

# Accuracy
my.statistics(test_binary$label, knn_pred)

```

## k = 5
```{r}
# k = 5 on Binary data
knn_pred <- knn(train_binary[,c("x","y")],
                test_binary[,c("x", "y")],
                cl = train_binary$label, 
                k = 5)

# Accuracy
my.statistics(test_binary$label, knn_pred)$accuracy

```

## k = 10
```{r}
# k = 10 on Binary data
knn_pred <- knn(train_binary[,c("x","y")],
                test_binary[,c("x", "y")],
                cl = train_binary$label, 
                k = 10)

# Accuracy
my.statistics(test_binary$label, knn_pred)$accuracy
```

## k = 15
```{r}
# k = 15 on Binary data
knn_pred <- knn(train_binary[,c("x","y")],
                test_binary[,c("x", "y")],
                cl = train_binary$label, 
                k = 15)

# Accuracy
my.statistics(test_binary$label, knn_pred)$accuracy
```

## k = 20
```{r}
# k = 20 on Binary data
knn_pred <- knn(train_binary[,c("x","y")],
                test_binary[,c("x", "y")],
                cl = train_binary$label, 
                k = 20)

# Accuracy
my.statistics(test_binary$label, knn_pred)$accuracy
```

## k = 25
```{r}
# k = 25 on Binary data
knn_pred <- knn(train_binary[,c("x","y")],
                test_binary[,c("x", "y")],
                cl = train_binary$label, 
                k = 25)

# Accuracy
my.statistics(test_binary$label, knn_pred)$accuracy
```



# KNN for Trinary classifier
## k = 3
```{r}
knn_pred <- knn(train_trinary[, c("x", "y")],
                test_trinary[,c("x","y")], 
                cl = train_trinary$label,
                k = 3)

my.statistics(test_trinary$label, knn_pred)$accuracy

```

## k = 5
```{r}
knn_pred <- knn(train_trinary[, c("x", "y")],
                test_trinary[,c("x","y")], 
                cl = train_trinary$label,
                k = 5)

my.statistics(test_trinary$label, knn_pred)$accuracy

```


