---
title: "Binary classifier regression model"
author: "Yograj Karki"
date: "5/19/2021"
output:
  pdf_document: default
  word_document: default
---

# Loading libraries
```{r warning = F}
library(caTools)
```


# Loading the dataset
Here I created csv file by copying the contents from the arff file, and then fixed the attributes of certain columns.
```{r}
setwd("~/MSDS/DSC520/dsc520/Binary_classifier")

data <- read.csv("binary-classifier-data.csv")
str(data)
```

# Splitting the data into train and test datasets
```{r}
# Splitting the data into train and test with the 80:20 ratio
dt = sort(sample(nrow(data), nrow(data)*0.8))

# Training dataset
train<-data[dt,]

# Testing dataset
test<-data[-dt,]
```


# Fitting the data to a logistic regression model

Here I used the glm() function to fit the model and the method was binomial logistic regression. 
```{r}
# fitting the model
logistic <- glm(label ~ x+y, data = train, family="binomial")

# summary of the model
summary(logistic)
```


# Checking the accuracy of the model
Running the test data throught the model
```{r echo = TRUE}
res <- predict(logistic, test, type="response")
res
```

# Confusion matrix
```{r}
confmatrix <- table(Actual_value = test$label, Predicted_value = res >0.5)
confmatrix
```

# Accuracy percentage
```{r}
#calculating accuracy percentage
(confmatrix[[1]]+confmatrix[[2]])/sum(confmatrix)
```
Our model turned out to be only 52% accurate.

